Blame and Coercion: Together Again for the Second Time

### 2019-10-17 paper.tex
comments

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. The rules are tricky...
  some words don't have a plural form, like ``blame''. -Jeremy
  \\---\\
  Will fix grammar after fixing all other problems -KC}
\todo[inline]{The first paragraph of the introduction should summarize
  the entire introduction, and in some sense, the entire paper. -Jeremy}

@@

\citet{siek2009exploring} present another intermediate langauge for gradually 
typed language, cast calculus \castCalculus\ and achieve both blame-tracking 
and space-efficiency for this language by decorating \citet{herman2010space}'s 
coercions with labels. \castCalculus\ is simply typed lambda calculus extended 
with the dynamic type and cast expressions. The labels in \castCalculus{}, 
unlike their counterparts in the Blame Calculus, have no polarity. 
They also propose that there are four blame strategies for \castCalculus{}:
$ \{Lazy, Eager\} \times \{D, UD\} $. Eager strategies detect more potential 
type errors, but blame more programs. $ D $ and $ UD $ assign blame labels 
differently. They claim that their \lazyUD\ cast calculus is the same as the 
Blame Calculus, modulo label polarity, and prove that their their \lazyUD\ 
coercion calculus simulates their \lazyUD{} cast calculus. It is unknown, 
however, whether the simulation in the other direction also works.
\todo[inline]{Is it important to this paper whether there is a
simulation in the other direction? -Jeremy
\\ --- \\
I guessed to say coercion is correct the other direction is 
necessary. -KC}

@@

Secondly, the definition of normal coercion lies unnecessarily on top 
of coercion: all coercions are normal when they are initially constructed, and 
\citet{siek2012interpretations} have shown an function that 
composes and produces normal coercions.
%
\todo[inline]{Many researchers would view this as a positive thing,
  not a negative. For example, the notion of normal forms comes up
  over and over again in mathematics and logic. -Jeremy
\\ --- \\
I guess having a notion of normal forms is more convenient when a 
notion of normalization is convenient. But we don't even need a notion of 
normalization for cast representations. -KC}

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hypercoercion calculus 
bisimulates the \lazyD{} (resp. \lazyUD{}) cast calculus. This is 
the first bisimulation proof for the \lazyD{} cast calculus as far as we know.
We care more about lazy strategies because recently, \citet{new2019gradual} 
show that \eager{} strategies are incompatible with $\eta$-equivalence of 
functions, which suggest that these strategies are not very ideal. 
\todo[inline]{Explain why we care about \lazyD{} and \lazyUD{}. -Jeremy
  \\ --- \\ Explained, not sure if it suffice. -KC
  \\ --- \\
  That's good that you explained why lazy instead of eager.
  (Also, you should add that we are interested in lazy because
  that's what Grift does.)\\
  But you haven't explained why we are interested in both D and UD. -JS
}

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ cast 
calculi. 
Section \ref{sec:coercion-calculus} reviews the state-of-art of $Lazy$ Coercion 
Calculi. 
In section \ref{sec:hypercoercion} we present hypercoercion.
Section \ref{sec:conclude} concludes.
\todo[inline]{Lots of missing sections. -Jeremy}

@@

\figref{fig:blame-static} defines the syntax of cast calculi and its
static semantics. We extend the syntax in \citet{siek2009exploring} with sum 
types and product types. Besides, we annotate function codomain explicitly 
because we need the codomain when evaluating cast calculus with hypercoercions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{The term typing rule for casts is broken. -Jeremy
\\ --- \\ Fixed -KC}
\todo[inline]{Why do you annotate the codomain of lambda abstractions
	explicitly? -Jeremy
	\\---\\
	Because I apply $ id(\cdot) $ to the codomain in the HC CEK. -KC
        \\---\\
        The answer you put in the text is to vague. -JS
        }
Let $ T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of cast calculi as a 
CEK machine \citep{felleisen1986control}. CEK machines include continuation as 
part of their states. They evaluate expressions with two interleaving 
processes: 
(1) dive into an expression and extend the continuation;
(2) return a value to the current continuation.
\todo[inline]{Briefly present the main ideas regarding how a CEK
	machine works. (2-3 sentences) -Jeremy
\\---\\Fixed -KC}

Most existing proofs about cast calculi defines the dynamic semantics with 
reduction semantics. We use a CEK machine instead because one of us is more 
familiar with CEK machines, and that he believes CEK machines are more 
convenient because space-efficiency is about compressing continuations.
\todo[inline]{Explain why you are using a CEK machine instead of
	reduction semantics. -Jeremy
\\---\\Fixed -KC}

\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
	the value first and the cast second. -Jeremy
	\\---\\
	Could you tell me why taking the value first is better?
	I am always taking the target of induction first. -KC
	\\---\\
        It's better because it matches the postfix syntax we are using for
        casts. Currently line 2 of applyCast for D is difficult to read,
        compare it with line 4 of applyCast for UD. -JS
        }

@@

The reduction relation $ \judgeDCreduce{s}{s} $ depends on $ applyCast $, a 
function depending on blame strategies. We will give the definitions of $ 
applyCast $ shortly.
To evaluate a cast 
expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one.
\todo[inline]{Change the rules for case to be small step, i.e.,
	the case function should not be recursive. -Jeremy}
Along the way, the machine accumulates the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

@@

% the following are temporary -Jeremy
\clearpage
\pagebreak

@@

\todo[inline]{Give a summary statement about the shape of the normal form. 
-Jeremy
\\---\\The three-part shape is not obvious before the transformation at 
the end of the next paragraph. So the summary is given after that. -KC}

@@

With the properties of surely \lazyD{} cast representation, it is 
straightforward to proof that 
all machines instantiated with these cast representations bisimulate each 
other. 
The state bisimilarity ($\eqvS{C}{C}{s}{s}$) is 
derived from the definition of states.
\todo[inline]{The term ``bisimilarity'' has a different technical meaning from
	``bisimulation relation''. I don't think you are using it properly
	in the above sentence. -Jeremy}
We omit subscripts when there is no 
ambiguity. \begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}

>>>>

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. The rules are tricky...
  some words don't have a plural form, like ``blame''. -JS
  \\---\\
  Will fix grammar after fixing all other problems -KC}
\todo[inline]{The first paragraph of the introduction should summarize
  the entire introduction, and in some sense, the entire paper. -JS}

@@

\citet{siek2009exploring} present another intermediate langauge for gradually 
typed language, cast calculus \castCalculus\ and achieve both blame-tracking 
and space-efficiency for this language by decorating \citet{herman2010space}'s 
coercions with labels. \castCalculus\ is simply typed lambda calculus extended 
with the dynamic type and cast expressions. The labels in \castCalculus{}, 
unlike their counterparts in the Blame Calculus, have no polarity. 
They also propose that there are four blame strategies for \castCalculus{}:
$ \{Lazy, Eager\} \times \{D, UD\} $. Eager strategies detect more potential 
type errors, but blame more programs. $ D $ and $ UD $ assign blame labels 
differently. They claim that their \lazyUD\ cast calculus is the same as the 
Blame Calculus, modulo label polarity, and prove that their their \lazyUD\ 
coercion calculus simulates their \lazyUD{} cast calculus. It is unknown, 
however, whether the simulation in the other direction also works.
\todo[inline]{Is it important to this paper whether there is a
simulation in the other direction? -JS
\\ --- \\
I guessed to say coercion is correct the other direction is 
necessary. -KC}

@@

Secondly, the definition of normal coercion lies unnecessarily on top 
of coercion: all coercions are normal when they are initially constructed, and 
\citet{siek2012interpretations} have shown an function that 
composes and produces normal coercions.
%
\todo[inline]{Many researchers would view this as a positive thing,
  not a negative. For example, the notion of normal forms comes up
  over and over again in mathematics and logic. -JS
\\ --- \\
I guess having a notion of normal forms is more convenient when a 
notion of normalization is convenient. But we don't even need a notion of 
normalization for cast representations. -KC
\\ --- \\
FWIW, I agree with you. But the point is that the argument is not
likely to be convincing to the reader, so we're probably better off
not wasting space on it. -JS
}

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hypercoercion calculus 
bisimulates the \lazyD{} (resp. \lazyUD{}) cast calculus. This is 
the first bisimulation proof for the \lazyD{} cast calculus as far as we know.
We care more about lazy strategies because recently, \citet{new2019gradual} 
show that \eager{} strategies are incompatible with $\eta$-equivalence of 
functions, which suggest that these strategies are not very ideal. 
\todo[inline]{Explain why we care about \lazyD{} and \lazyUD{}. -JS
  \\ --- \\ Explained, not sure if it suffice. -KC
  \\ --- \\
  That's good that you explained why lazy instead of eager.
  (Also, you should add that we are interested in lazy because
  that's what Grift does.)\\
  But you haven't explained why we are interested in both D and UD. -JS
}

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ cast 
calculi. 
Section \ref{sec:coercion-calculus} reviews the state-of-art of $Lazy$ Coercion 
Calculi. 
In section \ref{sec:hypercoercion} we present hypercoercion.
Section \ref{sec:conclude} concludes.
\todo[inline]{Lots of missing sections. -JS}

@@

\figref{fig:blame-static} defines the syntax of cast calculi and its
static semantics. We extend the syntax in \citet{siek2009exploring} with sum 
types and product types. Besides, we annotate function codomain explicitly 
because we need the codomain when evaluating cast calculus with hypercoercions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -JS
\\ --- \\ Fixed -KC}
\todo[inline]{The term typing rule for casts is broken. -JS
\\ --- \\ Fixed -KC}
\todo[inline]{Why do you annotate the codomain of lambda abstractions
	explicitly? -JS
	\\---\\
	Because I apply $ id(\cdot) $ to the codomain in the HC CEK. -KC
        \\---\\
        The answer you put in the text is too vague. -JS
        }
Let $ T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of cast calculi as a 
CEK machine \citep{felleisen1986control}. CEK machines include continuation as 
part of their states. They evaluate expressions with two interleaving 
processes: 
(1) dive into an expression and extend the continuation;
(2) return a value to the current continuation.
\todo[inline]{Briefly present the main ideas regarding how a CEK
	machine works. (2-3 sentences) -JS
\\---\\Fixed -KC}

Most existing proofs about cast calculi defines the dynamic semantics with 
reduction semantics. We use a CEK machine instead because one of us is more 
familiar with CEK machines, and that he believes CEK machines are more 
convenient because space-efficiency is about compressing continuations.
\todo[inline]{Explain why you are using a CEK machine instead of
	reduction semantics. -JS
        \\---\\Fixed -KC
        \\---\\
        Change ``one of us'' to ``first author''. -JS
}

\todo[inline]{Please fill in the missing reduction rules. -JS}
\todo[inline]{Change the order of parameters for applyCast to take
	the value first and the cast second. -JS
	\\---\\
	Could you tell me why taking the value first is better?
	I am always taking the target of induction first. -KC
	\\---\\
        It's better because it matches the postfix syntax we are using for
        casts. Currently line 2 of applyCast for D is difficult to read,
        compare it with line 4 of applyCast for UD. -JS
        }

@@

The reduction relation $ \judgeDCreduce{s}{s} $ depends on $ applyCast $, a 
function depending on blame strategies. We will give the definitions of $ 
applyCast $ shortly.
To evaluate a cast 
expression, the machine move the 
cast to the continuation and evaluate the inner expression. To apply a casted 
function, the machine firstly cast $ v_1 $, the operand, then apply the casted 
operand to $ v_2 $, the un-casted function, and finally cast the result of the 
function application. To take out the first (resp. second) part of a casted 
pair, the machine firstly take out the first (resp. second) part of $ v $, the 
un-casted pair, and cast the result. To case-split a sum, the machine looks 
deep inside the value to find out whether it is a left injection or a right 
one.
\todo[inline]{Change the rules for case to be small step, i.e.,
	the case function should not be recursive. -JS}
Along the way, the machine accumulates the pending casts onto the 
continuations. To cast a value, the machine try to apply the cast to the value. 
If the cast succeeds, the machine returns the result to the next continuation. 
If the cast fails, the machine halts with the blame label.

@@

% the following are temporary -JS
\clearpage
\pagebreak

@@

\todo[inline]{Give a summary statement about the shape of the normal form. 
-JS
\\---\\The three-part shape is not obvious before the transformation at 
the end of the next paragraph. So the summary is given after that. -KC}

@@

With the properties of surely \lazyD{} cast representation, it is 
straightforward to proof that 
all machines instantiated with these cast representations bisimulate each 
other. 
The state bisimilarity ($\eqvS{C}{C}{s}{s}$) is 
derived from the definition of states.
\todo[inline]{The term ``bisimilarity'' has a different technical meaning from
	``bisimulation relation''. I don't think you are using it properly
	in the above sentence. -JS}
We omit subscripts when there is no 
ambiguity. \begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}
